Manuscript Figures for Genome Biology

### 2018-07-23 manuscript_figures/manuscript_figures.Rmd
Add impact of covariates averaged over datasets in casestudies

# regexp to construct master data table separating case study datasets, methods
# and covariates
obj.names <- str_split(objects, pattern = "/")
casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[1]]))
extD <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 2)]]))
extD <- gsub("samplesize", "n", extD)
extM <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 3)]]))
extC <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 4)]]))
inform <-  ifelse(grepl("uninf", objects), "uninf", "inf")
cov <- inform
df <- data.frame(filename, casestudy, dataset, inform, cov, stringsAsFactors = FALSE) %>%
  mutate(dataset = ifelse(grepl("enigma", dataset), paste(dataset, extD, sep="-"),
                          dataset)) %>%  
  mutate(dataset = ifelse(casestudy == "scRNAseq", paste(dataset, extM, sep="-"),
                          dataset)) %>%
  mutate(cov = ifelse(casestudy == "scRNAseq" & inform != "uninf", extC,  cov)) %>%
  mutate(cov = ifelse(casestudy == "ChIPseq" & inform != "uninf", "region-width",  cov)) %>%
  mutate(cov = ifelse(casestudy == "GSEA" & inform != "uninf", "set-size",  cov)) %>%
  mutate(cov = ifelse(casestudy == "GWAS" & inform != "uninf", extD,  cov)) %>%
  mutate(cov = ifelse(casestudy == "microbiome" & inform != "uninf", 
                      ifelse(grepl("abun", filename), "abundance", "ubiquity"),  cov))%>%
  mutate(cov = ifelse(casestudy == "RNAseq" & inform != "uninf", "mean",  cov))

tab <- tidy_df(objects=objects, colLabels=filename, 
              fill="meanRank", annotate=NULL, alpha =0.10) %>%
  rename(casestudy = "filename") %>%
  mutate(filename = gsub(".rds", "", filename))

dftab <- left_join(tab, df, by="filename") %>%
  filter(method %in% cov.methods) %>%
  select(filename, casestudy, dataset, method, cov, inform, nrejects) %>%
  mutate(id = paste(casestudy, dataset, method, sep = "-"))

uninftab <- dftab %>% 
  filter(inform == "uninf") %>%
  select(id, nrejects) %>%
  rename(nrejects = "uninf")

@@

dftab <- left_join(inftab, uninftab, by = "id") %>%
  mutate(diff_inf = nrejects - uninf) %>%
  mutate(diff_inf_percent = diff_inf / nrejects) %>%
  mutate(method = factor(method, levels = rowOrder)) %>%
  group_by(casestudy, cov) %>%
  complete(method, dataset)

@@

>>>>

# regexp to construct master data table separating case study datasets, methods
# and covariates
obj.names <- str_split(objects, pattern = "/")
casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[1]]))
extD <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 2)]]))
extD <- gsub("samplesize", "n", extD)
extM <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 3)]]))
extC <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 4)]]))
inform <-  ifelse(grepl("uninf", objects), "uninf", "inf")
cov <- inform
df <- data.frame(filename, casestudy, dataset, inform, cov, stringsAsFactors = FALSE) %>%
  mutate(dataset = ifelse(grepl("enigma", dataset), paste(dataset, extD, sep="-"),
                          dataset)) %>%  
  mutate(dataset = ifelse(casestudy == "scRNAseq", paste(dataset, extM, sep="-"),
                          dataset)) %>%
  mutate(cov = ifelse(casestudy == "scRNAseq" & inform != "uninf", extC,  cov)) %>%
  mutate(cov = ifelse(casestudy == "ChIPseq" & inform != "uninf", "region-width",  cov)) %>%
  mutate(cov = ifelse(casestudy == "GSEA" & inform != "uninf", "set-size",  cov)) %>%
  mutate(cov = ifelse(casestudy == "GWAS" & inform != "uninf", extD,  cov)) %>%
  mutate(cov = ifelse(casestudy == "microbiome" & inform != "uninf", 
                      ifelse(grepl("abun", filename), "abundance", "ubiquity"),  cov))%>%
  mutate(cov = ifelse(casestudy == "RNAseq" & inform != "uninf", "mean",  cov)) %>%  
  mutate(cov = ifelse(cov == "det", "detection-rate", cov))

tab <- tidy_df(objects=objects, colLabels=filename, 
              fill="meanRank", annotate=NULL, alpha =0.10) %>%
  rename(filename = "casestudy") %>%
  #rename(casestudy = "filename") %>%
  mutate(filename = gsub(".rds", "", filename))

dftab <- left_join(tab, df, by="filename") %>%
  dplyr::filter(method %in% cov.methods) %>%
  dplyr::select(filename, casestudy, dataset, method, cov, inform, nrejects) %>%
  dplyr::mutate(id = paste(casestudy, dataset, method, sep = "-"))

uninftab <- dftab %>% 
  dplyr::filter(inform == "uninf") %>%
  dplyr::select(id, nrejects) %>%
  dplyr::rename("uninf" = "nrejects")
  #dplyr::rename(nrejects = "uninf")

inftab <- dftab %>%
  dplyr::filter(inform == "inf") %>%
  dplyr::select(-inform)

@@

dftab <- left_join(inftab, uninftab, by = "id") %>%
  dplyr::mutate(diff_inf = nrejects - uninf) %>%
  dplyr::mutate(diff_inf_percent = diff_inf / nrejects) %>%
  dplyr::mutate(method = factor(method, levels = rowOrder)) %>%
  dplyr::group_by(casestudy, cov) %>%
  complete(method, dataset)

@@

Make the Figure above, but average over all datasets within a case study.

```{r}
dftab <- dftab %>%
  group_by(casestudy, method) %>%
  summarize(diff_inf_percent = mean(diff_inf_percent, na.rm = TRUE))

maxVals <- dftab %>% 
  group_by(casestudy) %>%
  summarize(top = max(diff_inf_percent, na.rm=TRUE))

dftab <- left_join(dftab, maxVals, by = c("casestudy")) %>%
  mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA))

Fig_impactofinfcov <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_percent)) + 
        geom_tile() + 
        scale_fill_gradientn("% Increase\nin rejections",
                             labels = scales::percent, trans="log1p",
                             colours = c("darkblue","white","darkred"), 
                             values = scales::rescale(c(-.05,0,1)),
                             guide = "colorbar", limits=c(-.05,1)) +
        theme_bw() +
        xlab("Case Study") +
        ylab("Method") +
        ggtitle("Impact of informative covariate on number of rejections") +
        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
        geom_text(aes(label = ifelse(is.na(top), "", scales::percent(top))),
                  color = "black", size = 2)

Fig_impactofinfcov
ggsave(file.path(outdir, "Figure_impactofinfcov_avg.pdf"), width = 5.75, height = 4)
```
